---
title: "Pathway based prediction of the *Caenorhabditis elegans* lipidome"
package: LipidNetworkPredictR
output:  
    BiocStyle::html_document:
        toc_float: true
vignette: >
  %\VignetteIndexEntry{lipid_prediction_reaction_order}
  %\VignettePackage{LipidNetworkPredictR-vignette}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)
```

This vignette describes the complete workflow for the prediction of
*Caenorhabditis elegans* (*C. elegans*) lipids performed by Witting et
al. (in preparation). All template reactions at the current stage have been 
curated from the WormJam model.

```{r setup}
# load required library
library(LipidNetworkPredictR)
library(dplyr)
library(rgoslin)
```

In order to store all the results a data frame for the reactions is
created. Lipids abbreviations are just appended.

```{r data structures, message=FALSE, warning=FALSE}
## create data frame to store reactions
lipid_reactions <- data.frame()

## create vector to store shorthand notations of generated lipids
##lipid_abbrev <- vector()
lipid_df <- tibble()
```

The following abbreviations are used (in alphabetical order):

-   `AcylDHAP`: acyl-dihydroxyacetonephosphate,
-   `AlkylDHAP`: dihydroxyacetonephosphate,
-   `C1P`: Ceramide-1-phosphates [SP0205],
-   `CDPDG`: cytidine diphosphate diacylglycerol,
-   `CER`: N-acylsphingosines (ceramides) [SP0201],
-   `CL`: Diacylglycerophosphoglycerophosphodiradylglycerols
    (cardiolipin) [GP1201],
-   `CoA`: Fatty acyl CoAs [FA0705],
-   `DG`: Diacylglycerols [GL0201],
-   `DGO`: 1-alkyl,2-acylglycerols
-   `DHCER`: N-acylsphinganines (dihydroceramides) [SP0202],
-   `DHSM`: Dihydrosphingomyelins [SP0301],
-   `FA`: Fatty acids [FA01],
-   `FATOH`: Fatty alcohols [FA05],
-   `GLCCER`: glycosylceramide [SP0501],
-   `GPNAE`: glycerophospho-N-acylethanolamine,
-   `LNAPE`: lyso-N-Acyl-phosphatidylethanolamine,
-   `LPA`: 1-acylglycerophosphates [GP1005],
-   `LPAO`: 1-alkylglycerophosphates [GP1006],
-   `LPCO`: 1-alkylglycerophosphocholine [GP0106],
-   `LPEP`: 1Z-alkenylglycerophosphocholine [GP0107],
-   `NAE`: N-Acyl-phosphatidylethanolamine
-   `NAPE`: N-Acyl-phosphatidylethanolamine,
-   `NAPE_sn1`: sn1 N-Acyl-phosphatidylethanolamine
-   `NAPEP_sn1`: ...
-   `NAPE_sn2`: sn2 N-Acyl-phosphatidylethanolamine
-   `NAPEP_sn2`: ...
-   `NAPEO`: ...
-   `PA`: Diacylglycerophosphates [GP1001],
-   `PAO`: 1-alkyl,2-acylglycerophosphates [GP1002]
-   `PC`: Diacylglycerophosphocholine [GP0101],
-   `PCO`: 1-alkyl,2-acylglycerophosphocholine [GP0102]
-   `PE`: Diacylglycerophosphoethanolamines [GP0201],
-   `PEP`: 1(1Z-alkenyl),2-acylglycerophosphoethanolamines [GP0203],
-   `PEO`: 1-alkyl,2-acylglyerophosphoethanolamines [GP0202]
-   `PG`: Diacylglycerophosphoglycerols [GP0401],
-   `PGP`: Diacylglyerophosphoglycerophosphates [GP0501]
-   `PI`: Diacylglycerophosphoinositols [GP0601],
-   `PNAE`: N-acyl ethanolamine phosphates,
-   `PS`: Diacylglycerophosphoserines [GP0301],
-   `SM`: Ceramide phosphocholines (sphingomyelins) [SP0301],
-   `sn1LPC`: 1-acylglycerophosphocholine [GP0105],
-   `sn1LPE`: 1-acylglycerophosphoethanolamines [GP0205],
-   `sn1MG`: 1-acylglycerols [GL0101],
-   `sn2LPC`: 2-acylglycerophosphocholine [GP0105],
-   `sn2LPE`: 2-acylglycerophosphoethanolamines [GP0205],
-   `sn2MG`: 2-acylglycerols [GL0101]
-   `SPHINGA`: Sphinganines [SP0102],
-   `SPHINGO`: Sphing-4-enines (sphingosines) [SP0101],
-   `TG`: Triacylglycerols [GL0301]

## Fatty acids used for biosynthesis

First input for the prediction are fatty acids that were shown to be
present in *C. elegans* by different groups. A consensus list was
generated based on fatty acids from WormJam and detected in *C. elegans*.
A data frame of these fatty acids is supplied with this
package. Fatty acids used in glycero- and glycerophospholipids are
different from the ones used in sphingolipid biosynthesis. The pools are
only partially overlapping between the two. Therefore, two different
files for the two different pools are supplied. The following lines read
the fatty acids into a data frame and isolates the shorthand notation of
the fatty acids, which are further used for downstream processing.

```{r fatty acids 1, message=FALSE, warning=FALSE}
## read data frame with fatty acids detected in C. elegans
f <- system.file("extdata", "celegans_fatty_acids_GL_GP.txt",
    package = "LipidNetworkPredictR")

fatty_acid_gl_gp_df <- read.table(f, sep = "\t", header = TRUE, 
        comment.char = "", stringsAsFactors = FALSE) |> 
    data.frame()

## get short hand notations for fatty acids
FA_GPGL <- fatty_acid_gl_gp_df$shorthand

## take only a subset of those to speed things up
FA_GPGL <- FA_GPGL[FA_GPGL %in%
    c("FA(12:0)", "FA(14:0)", "FA(15:0)", "FA(16:0)", "FA(17:0)", "FA(18:0)", "FA(19:0)", "FA(20:0)" )]
lipid_info <- rgoslin::parseLipidNames(FA_GPGL)

## add lipids to data frame
lipid_df <- dplyr::bind_rows(lipid_df, lipid_info)
```

The lines below perform the same action for fatty acids used in
sphingolipid biosynthesis.

```{r fatty acids 2, message=FALSE, warning=FALSE}
## read data frame with fatty acids detected in C. elegans
f <- system.file("extdata", "celegans_fatty_acids_SPH.txt",
    package = "LipidNetworkPredictR")

fatty_acid_sph_df <- read.table(f, sep = "\t", header = TRUE,
    comment.char = "", stringsAsFactors = FALSE) |>
    data.frame()

## get short hand notations for fatty acids
FA_SPH <- fatty_acid_sph_df$shorthand

## take only a subset of those to speed things up
FA_SPH <- FA_SPH[FA_SPH %in%
    c("FA(12:0)", "FA(14:0)", "FA(15:0)", "FA(16:0)", "FA(17:0)", "FA(18:0)", "FA(19:0)", "FA(20:0)" )]
lipid_info <- rgoslin::parseLipidNames(FA_SPH)

## add lipids to data frame
lipid_df <- dplyr::bind_rows(lipid_df, lipid_info)
```

# GPGL pool

## Block 0: Preparation of Acyl-CoA

### Biosynthesis of CoA from FA (RHEA:15421)

Acyl groups are transfered in their activate form of acyl CoA (CoA).
Their synthesis requires ATP. *C. elegans* harbors different Acyl-CoA
synthetases and their substrate specificity is not established so far.
In order to be able to use all defined acyls they are converted to acyl
CoA. The reaction converts fatty acids (FA) to acyl-CoA. 

## Block 1: Biosynthesis and Degradation of Phosphaditic acid and Diacylglycerol

Phosphaditic acid (PA) and Diacylgylcerols (DG) are an important
intermediate for the biosynthesis of different lipids. The represent a
branching point between triacylglycerols and glycerophospholipids. They
are synthesized from lysosphosphaditic and phosphaditic acid by
successive acylation of glycerol 3-phosphate. The first block of
reactions handles the generation of PA and DG as well as their
degradation via monoacylglycerols to fatty acids.

### Biosynthesis of LPA (RHEA:15325)

The first step to obtain DG is the biosynthesis of lysophophaditic acid
from acyl CoA_GPGL and glycerol 3-phosphate. This reaction takes place
in the ER and the mitochondria. In the ER the reaction is catalyzed by
acl-5 (WBGene00011106) or acl-6 (WBGene00017261), while in the
mitochondria acl-4 (WBGene00018657) is used. No subcellular specificity
is currently used, therefore the assignments are merged together to
yield a single reaction.

### Biosynthesis of PA (RHEA:19709)

In the next step a second acyl group is transferred to the LPA to
produce PA. This reaction is catalyzed by acl-9 (WBGene00022646) or
acl-8 (WBGene00020264) or mboa-3 (WBGene00016934) or acl-1
(WBGene00010339) or acl-2 (WBGene00011543) in the ER. For the
mitochondrial counterpart no genes are known so far. acl-3
(WBGene00006491) might be a good candidate since it is predicted to
locate in the mitochondria.

### Biosynthesis of DG (RHEA:27429)

Diacylglycerols (DG) are produced from PA by dephosphorylation. DG are
important intermediates for biosynthesis of glycerophospholipids or
triacylglycerols (DG). A prominent enzyme catalyzing this reactions is
LPIN-1. Plasma membrane or ER WBGene00008749 or WBGene00018756 or
WBGene00020895 or WBGene00010425 nucleus WBGene00010425

### DG Kinase (RHEA:10272)

DG can be phosphorylated to create PA, which is the reversal of the
reaction above. Under normal conditions this reactions is inactive, but can be
activated yielding PA. Normally, DG is used for synthesis of TGs and phospho-
lipids, but phosphorylation yields a switch in DG to PA, which both have also
different signaling functions. In this case a lipid, which has been generated 
previously by a different reactions is produced. 

### Degradation of DG (RHEA:44712, RHEA:33275, sn2mg_to_sn1mg)

#### Production of 1-MG (RHEA:44712)

DG are degraded by the activity of different lipases yielding mono
acylglycerols. In the case of this reaction the fatty acid at sn1
position is removed yieling an sn2-MG. sn2-MG can be converted to
sn1-MG, which in turn can be re-acylated at a later stage.


#### Production of 2-MG (RHEA:33275)

DG are degraded by the activity of different lipases yielding mono 
acylglycerols. In the case of this reaction the fatty acid at sn1 position is 
removed yieling an sn2-MG. sn2-MG can be converted to sn1-MG, which in turn 
can be re-acylated at a later stage.

#### Acyl chain migration (sn2mg_to_sn1mg)

Acyl chain migration can happen to produced 1-Acyl-MGs from 2-Acyl-MGs. Both 
species are in equilibrium to each other. However, 1-Acyl-MGs can be reacylated
to yields DGs.

### Degradation of MG (RHEA:34019, RHEA:32871)

Both, sn1- and sn2-MG, can degraded to free fatty acids and glycerol,
which can be used for energy production upon conversion to Acyl-CoAs and import
into mitochondria for beta-oxidation. Other possibilities include the re-use in
lipid biosynthesis.

### Phosphorylation and reacylation of MG (RHEA:38463, RHEA:32947, RHEA:33747)

Similar to the degradation, all MG can be reacylated to yields DG again.

Additionally, 1-Acyl-MG can be also phosphorylated to yield LPA, which is then 
fed into lipid biosynthesis.

## Block 2: Biosynthesis and degradation of TG

Triacylglycerols (TG) are storage lipids in *C. elegans* and other
organisms. They are stored in lipid droplets. In *C. elegans* they are
also found in yolk for the production of eggs.

### Biosynthesis of TG (RHEA:10868)

TG are synthesized by acylation of DG in the ER and sequestered into lipid 
droplets. Since TGs harbor three acyl chains their combinatorial spaces is 
large.

### Degradation of TG (RHEA:33271)

TG are "symmetric" molecules, therefore from each TG two different DG
can produced.

## Block 3: Biosynthesis of Glycerophospholipids (PC, PE, PS)

The glycerophospholipids phosphatidylcholine, -ethanolamine and serine
are synthesized from DG by addition of headgroups. Additionally they can
be interconverted

### Biosynthesis of PC (RHEA:32939)

### Biosynthesis of PE (RHEA:32943)

### Biosynthesis of PS (RHEA:45088, RHEA:27606)

PS can be formed from PC (RHEA:45088) and from PE (RHEA:27606).

### Decarboxylation of PS to PE (RHEA:20828)

PE can be formed from PS species by decarboxylation. The reaction takes 
place only in mitochondria.

## Block 4: Remodeling and Degradation of Glycerophospholipids

### Phospholipase action on PC (RHEA:18689, RHEA:15801, RHEA:14445)

PLA1 cleaves of the acyl at the sn1 position producing an sn2LPC (RHEA:18689_.

PLA2 on PC (RHEA:15801).

PLD on PC (RHEA:14445).


### LPC to FA_GPGL (RHEA:15177, RHEA:44696)

sn1LPC to FA_GPGL.

sn2LPC to FA_GPGL.

### Reacylation of LPC to PC (Remodeling, RHEA:12937)

Only sn1LPC get re-acylated --\> constraining? = remodeling!

### Phospholipase action on PE (RHEA:44408)

PLA1 on PE (RHEA:44408).

PLA2 on PE (RHEA:44604).

PLD on PE (pe_to_pa).

### LPE to FA_GPGL (RHEA:32967)

sn1LPE to FA_GPGL (RHEA:32967).

sn2LPE to FA_GPGL (sn2lpe_to_fa)

### Reacylation of LPE to PE (Remodeling, RHEA:32995)

Only sn1LPE get re-acylated --\> constraining? = remodeling!

## Block 5: Biosynthesis of Glycerophospholipids (PG, PI, CL)

PG, PI and CL are synthesized based on the common precursor CDP-DG

### Biosynthesis of CDP-DG (RHEA:16229)

### Biosynthesis of PI (RHEA:11580)

### Biosynthesis of PGP and PG (RHEA:12593, RHEA:33751, RHEA:32931)

This reaction is taking place only in the mitochondria.

This reaction has no gene assignment at the moment (RHEA:33751).

The reaction from PG to CL is only taking place in mitochondria (RHEA:32931).

## Block 6: Biosynthesis of Etherlipids

### Generation of fatty alcohols (RHEA:52716)

Some etherlipids... mostly PE, small amounts in PC. This reaction is
constrained to 16:0, 18:0, 20:0!

### Generation of Acyl-DHAP (RHEA:17657)

### Generation of Alkyl-DHAP (RHEA:36171)

### Reduction of Alkyl-DHAP to LPA-O (RHEA:36175)

Enzyme not known so far in C. elegans!!! EC 1.1.1.101

### Biosynthesis of PA-O (RHEA:36235)

Same genes as for LPA to PA --\> substrate specificity not known yet.

### Biosynthesis of DG-O (RHEA:36239)

### Biosnythesis of PE-O (RHEA:36187)

Headgroup transfer via cept-1 and cept-2.

### Biosynthesis of PE-P (RHEA:22956)

reductase is unknown so far.

### Biosynthesis of PC-O (RHEA:36179)

same as for PC!

## Block 7: Remodeling of Etherlipids

### PLA2 on PE-O (peo_to_lpeo)

### Reacylation of PE-O (lpeo_to_peo)

### PLA2 on PE-P (pep_to_lpep)

### Reacylation of PE-P (lpep_to_pep)

### PLA2 on PC-O (pco_to_lpco)

### Remodeling of PC-O???

### Degradation of LPEP

## Block Y: Biosynthesis and Degradation of N-Acyl-Ethanolamines

This biosynthetic pathway is putative. Only a few enzymes have been
found so far.

### Biosynthesis of N-Acyl-phosphatidylethanolamine (NAPE) from PE and PC (pe_to_nape_sn1, peo_to_napeo_sn1, pep_to_napep_sn1, pe_to_nape_sn2, peo_to_napeo_sn2, pep_to_napep_sn2)

N-acyl-phosphatidylethanolamines are produced by the transfer of an acyl
group from PC to the nitrogen of PE. In *C. elegans* this reaction is so
far not known. It is unknown if the sn1 or sn2 position is transferred.
Since *C. elegans* contains about equal amounts of PE, PE-O and PE-P it
might be possible that the acyl group is transferred to all of them
producing NAPE, NAPE-O and NAPE-P.

transfer of Sn1 fatty acyl to PE (pe_to_nape_sn1)

transfer of Sn1 fatty acyl to PE-O (peo_to_napeo_sn1)

transfer of Sn1 fatty acyl to PE-P (pep_to_napep_sn1)

transfer of Sn2 fatty acyl to PE (pe_to_nape_sn2)

transfer of Sn2 fatty acyl to PE-O (peo_to_napeo_sn2)

transfer of Sn2 fatty acyl to PE-P (pep_to_napep_sn2)


### Biosynthesis of N-Acyl-ethanolamine (nape_to_nae, napeo_to_nae)

From NAPE (nape_to_nae)

From NAPEO (napeo_to_nae)

### Biosynthesis of from NAPE (nape_to_pnae)

From NAPE.

### Generation of N-Acyl-Ethanolamines from N-Acyl-Phosphoethanolamines

### Generation of LysoNAPE from NAPE (nape_to_lnape)

### Generation of GPNAE from LysoNAPE (RHEA:45420)

### Degradation of N-Acyl-Ethanolamines (nae_to_fa)

## Run the function 

```{r, message=FALSE, warning=FALSE}
reactions <- rbind(
    ## Block 0
    c(1, "RHEA:15421", "M_ATP + M_CoA + M_FA = M_PPi + M_AMP + M_AcylCoA"), ## FA_GPGL + CoA = PPi + AMP + CoA_GPGL
    ## Block 1
    c(2, "RHEA:15325", "M_Glycerol-3-P + M_AcylCoA = M_CoA + M_LPA"), ## G3P + CoA_GPGL = CoA + LPA
    c(3, "RHEA:19709", "M_LPA + M_AcylCoA = M_CoA + M_PA"), ## LPA + CoA_GPGL = CoA + PA
    c(4, "RHEA:27429", "M_H2O + M_PA = M_Pi + M_1,2-DG"), ## H2O + PA = Pi + DG
    c(5, "RHEA:10272", "M_ATP + M_1,2-DG = M_H+ + M_ADP + M_PA"), ## ATP + DG = H+ + ADP + PA
    c(6, "RHEA:44712", "M_H2O + M_1,2-DG = M_H+ + M_1-MG + M_FA"), ## H2O + DG = H+ + sn1MG + FA
    c(7, "RHEA:33275", "M_H2O + M_1,2-DG = M_H+ + M_2-MG + M_FA"), ## H2O + DG = H+ + sn2MG + FA
    c(8, "sn2mg_to_sn1mg", "M_2-MG <=> M_1-MG"), ## sn1MG <=> sn2MG
    c(9, "RHEA:34019", "M_H2O + M_1-MG = M_Glycerol + M_H+ + M_FA"), ## H2O + sn1MG = Glycerol + H+ + FA
    c(10, "RHEA:32871", "M_H2O + M_2-MG = M_Glycerol + M_H+ + M_FA"), ## H2O + sn2MG = Glycerol + H+ + FA
    c(11, "RHEA:38463", "M_1-MG + M_AcylCoA = M_CoA + M_1,2-DG"), ## sn1MG + CoA_GPGL = CoA + DG
    c(12, "RHEA:32947", "M_AcylCoA + M_1-MG = M_CoA + M_1,2-DG"), #################### CoA_GPGL + sn2MG = CoA + DG #################### should be 2-MG???
    c(13, "RHEA:33747", "M_ATP + M_1-MG = M_H+ + M_ADP + M_LPA"), ## ATP + 1-MG = H+ + ADP + LPA
    ## Block 2
    c(14, "RHEA:10868", "M_AcylCoA + M_1,2-DG = M_CoA + M_TG"), ## CoA_GPGL + DG = CoA + TG
    c(15, "RHEA:33271", "M_H2O + M_TG = M_H+ + M_FA + M_1,2-DG"), ## H2O + TG = H+ + FA + DG
    ## Block 3
    c(16, "RHEA:32939", "M_CDP-Choline + M_1,2-DG = M_H+ + M_CMP + M_PC"), ## CDP-Choline + DG = H+ + CMP + PC
    c(17, "RHEA:32943", "M_CDP-Ethanolamine + M_1,2-DG = M_H+ + M_CMP + M_PE"), ## CDP-Ethanolamine + DG = H+ + CMP + PE
    c(18, "RHEA:45088", "M_L-Serine + M_PC = M_Choline + M_PS"), ## L-Serine + PC = Choline + PS
    c(19, "RHEA:27606", "M_L-Serine + M_PE = M_Ethanolamine + M_PS"), ## L-Serine + PE = Ethanolamine + PS
    c(20, "RHEA:20828", "M_H+ + M_PS = M_CO2 + M_PE"), ## H+ + PS = CO2 + PE
    ## Block 4
    c(21, "RHEA:18689", "M_H2O_c + M_PC_c = M_2-LPC_c + M_FA_c"), ## H2O + PC = sn2LPC + FA_GPGL
    c(22, "RHEA:15801", "M_H2O + M_PC = M_1-LPC + M_FA"), ## H2O + PC = sn1LPC + FA_GPGL
    c(23, "RHEA:14445", "M_H2O + M_PC = M_Choline + M_PA"), ## H2O + PC = Choline + PA
    c(24, "RHEA:15177", "M_H2O + M_1-LPC = M_Glycerophosphocholine + M_H+ + M_FA"), ## H2O + sn1LPC = Glycerophosphocholine + H+ + FA_GPGL 
    c(25, "RHEA:44696", "M_H2O + M_2-LPC = M_Glycerophosphocholine + M_H+ + M_FA"), ## H2O + sn2LPC = Glycerophosphocholine + H+ + FA_GPGL
    c(26, "RHEA:12937", "M_1-LPC_c + M_AcylCoA_c = M_PC_c + M_CoA_c"), ## sn1LPC + CoA_GPGL = PC + CoA_GPGL
    c(27, "RHEA:44408", "M_H2O + M_PE = M_H+ + M_FA + M_2-LPE"), ## H2O + PE = H+ + FA_GPGL + sn2LPE
    c(28, "RHEA:44604", "M_H2O + M_PE = M_H+ + M_FA + M_1-LPE"), ## H2O + PE = H+ + FA_GPGL + sn1LPE
    c(29, "pe_to_pa", "M_H2O + M_PE <=> M_Ethanolamine + M_H+ + M_PA"), ## H2O + PE <=> Ethanolamine + H+ + PA
    c(30, "RHEA:32967", "M_H2O + M_1-LPE = M_H+ + M_FA + M_Glycerophosphoethanolamine"), ## H2O + sn1LPE = H+ + FA_GPGL + Glycerophosphoethanolamine
    c(31, "sn2lpe_to_fa", "M_H2O + M_2-LPE = M_H+ + M_FA + M_Glycerophosphoethanolamine"), ## H2O + sn2LPE = H+ + FA_GPGL + Glycerophosphoethanolamine
    c(32, "RHEA:32995", "M_AcylCoA + M_1-LPE = M_CoA + M_PE"), ## CoA_GPGL + sn1LPE = CoA + PE
    ## Block 5
    c(33, "RHEA:16229", "M_CTP + M_PA = M_PPi + M_CDP-DG"), ## CTP + PA = PPi + CDP-DG
    c(34, "RHEA:11580", "M_myo-Inositol + M_CDP-DG = M_H+ + M_CMP + M_PI"), ## myo-Inositol + CDP-DG = H+ + CMP + PI
    c(35, "RHEA:12593", "M_Glycerol-3-P + M_CDP-DG = M_H+ + M_CMP + M_PGP"), ## Glycerol-3-P + CDP-DG = H+ + CMP + PGP
    c(36, "RHEA:33751", "M_H2O + M_PGP = M_Pi + M_PG"), ## H2O + PGP = Pi + PG
    c(37, "RHEA:32931", "M_CDP-DG + M_PG = M_H+ + M_CMP + M_CL"), ## CDP-DG + PG = H+ + CMP + CL
    ## Block 6
    c(38, "RHEA:52716", "M_AcylCoA + 2 M_NADPH + 2 M_H+ = M_FATOH + 2 M_NADP + M_CoA"), ## AcylCoA_GPGL + 2 NADPH + 2 H+ = FATOH + 2 NADP + CoA
    c(39, "RHEA:17657", "M_Dihydroxyacetone-P + M_AcylCoA = M_CoA + M_AcylDHAP"), ## Dihydroxyacetone-P + CoA_GPGL = CoA + AcylDHAP 
    c(40, "RHEA:36171", "M_AcylDHAP + M_FATOH = M_H+ + M_FA + M_AlkylDHAP"), ## AcylDHAP + FATOH = H+ + FA_GPGL + M_AlkylDHAP
    c(41, "RHEA:36175", "M_H+ + M_NADPH + M_AlkylDHAP = M_LPA-O + M_NADP"), ## H+ + NADPH + AlkylDHAP = LPAO + NADP
    c(42, "RHEA:36235", "M_LPA-O + M_AcylCoA = M_PA-O + M_CoA"), ## LPAO + CoA_GPGL = PAO + CoA
    c(43, "RHEA:36239", "M_H2O + M_PA-O = M_Pi + M_DG-O"), ## H2O + PAO = Pi + DGO
    c(44, "RHEA:36187", "M_CDP-Ethanolamine + M_DG-O = M_H+ + M_CMP + M_PE-O"), ## CDP-Ethanolamine + DGO = H+ + CMP + PEO
    c(45, "RHEA:22956", "M_PE-O + M_Fe2+cytochrome_b5 + 2 M_H+ + M_O2 = M_PE-P + M_Fe3+cytochrome_b5 + 2 M_H2O"), ## PEO + Fe2+cytochrome_b5 + 2 H+ + O2 = PEP + Fe3+cytochrome_b5 + 2 H2O
    c(46, "RHEA:36179", "M_CDP-Choline + M_DG-O = M_H+ + M_CMP + M_PC-O"), ## CDP-Choline + DGO = H+ + CMP + PCO
    ## Block 7
    c(47, "peo_to_lpeo", "M_H2O + M_PE-O = M_H+ + M_LPE-O+ M_FA"), ## H2O + PEO = H+ + LPEO + FA_GPGL
    c(48, "lpeo_to_peo", "M_AcylCoA + M_LPE-O = M_CoA + M_PE-O"), ## CoA_GPGL + LPEO = CoA + M_PEO
    c(49, "pep_to_lpep", "M_H2O + M_PE-P = M_H+ + M_LPE-P + M_FA"), ## H2O + PEP = H+ + LPEP + FA
    c(50, "lpep_to_pep", "M_AcylCoA + M_LPE-P = M_CoA + M_PE-P"), ## CoA_GPGL + LPEP = CoA + PEP
    c(51, "pco_to_lpco", "M_H2O + M_PC-O = M_H+ + M_LPC-O + M_FA"), ## H2O + PCO = H+ + LPCO + FA_GPGL
    ## Block Y
    c(52, "pe_to_nape_sn1", "M_PE + M_PC = M_NAPE + M_2-LPC"), ## PE + PC = NAPE + sn2LPC"
    c(53, "peo_to_napeo_sn1", "M_PE-O + M_PC = M_NAPEO + M_2-LPC"), ## PEO + PC = NAPEO + sn2LPC
    c(54, "pep_to_napep_sn1", "M_PE-P + M_PC = M_NAPEP + M_2-LPC"), ## PEP + PC = NAPEP + sn2LPC
    c(55, "pe_to_nape_sn2", "M_PE + M_PC = M_NAPE + M_1-LPC"), ## PE + PC = NAPE + sn1LPC
    c(56, "peo_to_napeo_sn2", "M_PE-O + M_PC = M_NAPEO + M_1-LPC"), ## PEO + PC = NAPEO + sn1LPC
    c(57, "pep_to_napep_sn2", "M_PE-P + M_PC = M_NAPEP + M_1-LPC"), ## PEP + PC = NAPEP + sn1LPC
    c(58, "nape_to_nae", "M_NAPE + M_H2O <=> M_NAE + M_PA"), ## NAPE + H2O <=> NAE + PA
    c(59, "napeo_to_nae", "M_NAPEO + M_H2O <=> M_NAE + M_PA-O"), ## akac2nape + H2O <=> NAE + PAO
    c(60, "nape_to_pnae", "M_LNAPE + M_H2O <=> M_pnae + M_1,2-DG"), ## LNAPE + H2O <=> PNAE + DG
    c(61, "nape_to_lnape", "M_LNAPE + M_H2O <=> M_LNAPE + M_FA"), ## LNAPE + H2O <=> LNAPE + FA_GPGL
    c(62, "RHEA:45420", "M_LNAPE + M_H2O = M_GPNAE + M_FA"), ## LNAPE + H2O = GPNAE + FA_GPGL
    c(63, "nae_to_fa", "M_H2O + M_NAE = M_Ethanolamine + M_H+ + M_FA") ## H2O + nae = Ethanolamine + H+ + FA_GPGL
)

reactions <- data.frame(order = as.numeric(reactions[, 1]), RHEA = reactions[, 2],
    reactions = reactions[, 3])
 
## run the function
reactions_GPGL <- create_reactions(substrates = list(FA = FA_GPGL), reactions = reactions)
```


# SPH pool

## Block 0: Preparation of Acyl-CoA

### Biosynthesis from FA to CoA (RHEA:15421)

Likewise, the fatty acyls for the synthesis of sphingolipids are
required as acyl-CoA. The main differences are that fatty acyls in
sphingolipids are generally saturated and longer compared to glycero-
and glycerophospholipids and often contain a hydroxygroup at the second
position of the acyl.


## Block X: Biosynthesis and degradation of sphingolipids

### Biosynthesis of Dihydroceramides (sphinga_to_dhcer)

### Biosynthesis of Dihydrosphingomyelins (RHEA:44620)

### Degradation of Dihydrosphingomyelins (DHSM, RHEA:19253)

### Biosynthesis of Ceramides (CER, dhcer_to_cer)

### Biosynthesis of Sphingomyelins (SM, RHEA:18765)

### Degradation of Sphingomyelins (SM) to Ceramides (CER) (sm_to_cer)

### Biosynthesis of Ceramide-1-Phosphates (C1P) from Ceramides (CER) (RHEA:17929)

### Degradation of Ceramide-1-Phosphates (C1P) to Ceramides (CER) (c1p_to_cer)

Gene is unknown. 

### Biosynthesis of Glucosylceramides (RHEA:12088)


## Run the function
```{r, message=FALSE, warning=FALSE}
reactions <- rbind(
    c(1, "RHEA:15421", "M_ATP + M_CoA + M_FA = M_PPi + M_AMP + M_AcylCoA"), ## ATP + CoA + FA_SPH = PPi + AMP + CoA_SPH
    ## Block X
    c(2, "sphinga_to_dhcer", "M_AcylCoA + M_c17isosphgn <=> M_CoA + M_DHCER"), ## CoA_SPH + c17isosphgn <=> CoA + DHCER
    c(3, "RHEA:44620", "M_P-Choline + M_DHCER = M_1,2-DG + M_DHSM"), ## P-Choline + DHCER = DG + DHSM
    c(4, "RHEA:19253", "M_H2O + M_DHSM = M_P-Choline + M_H+ + M_DHCER"), ## H2O + DHSM = P-Choline + H+ + DHCER
    c(5, "dhcer_to_cer", "M_H+ + M_NADH + M_O2 + M_DHCER = 2 M_H2O + M_NAD + M_CER"), ## H+ + NADH + O2 + DHCER = 2 H2O + NAD + CER
    c(6, "RHEA:18765", "M_P-Choline + M_CER = M_1,2-DG + M_SM"), ## P-Choline + CER = 1,2-DG + SM
    c(7, "sm_to_cer", "M_H2O + M_SM = M_P-Choline + M_H+ + M_CER"), ## H2O + SM = P-Choline + H+ + CER
    c(8, "RHEA:17929", "M_ATP + M_CER = M_H+ + M_ADP + M_C1P"), ## ATP + CER = H+ + ADP + C1P
    c(9, "c1p_to_cer", "M_H2O + M_C1P = M_Pi + M_CER"), ## H2O + C1P = Pi + CER
    c(10, "RHEA:12088", "M_UDP-Glucose + M_CER = M_H+ + M_UDP + GLCCER") ## UDP-Glucose + CER = H+ + UDP + GLCCER
)

reactions <- data.frame(order = as.numeric(reactions[, 1]), RHEA = reactions[, 2],
    reactions = reactions[, 3])
 
## run the function
reactions_SPH <- create_reactions(substrates = list(FA = FA_SPH), reactions = reactions)
```


